---
title: "Combining hierarchical models and habitat data improves assessment of data limited salmon stocks "
subtitle:  |
    | Marine and Coastal Fisheries
    | Paper analysis
author: "W.I. Atlas, D.M. Glaser, B.M. Connors, C.A. Holt, D.A. Greenberg, D.T. Selbie, S. Cox-Rogers, C. Carr-Harris, E. Hertz, and J.W. Moore."
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output/html") })
output: 
  bookdown::html_document2:
    theme: cerulean
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: false
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: sentence
#bibliography: references.bib
#csl: "https://raw.githubusercontent.com/citation-style-language/styles/master/canadian-journal-of-fisheries-and-aquatic-sciences.csl"
link-citations: true
always_allow_html: yes
---

```{r pkg-setup, echo = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
library(here)
library(ggh4x) #to control facetting
library(scales) #for pretty_breaks()
library(rstan) #to interface R with Stan
library(bayesplot) #for mcmc_combo plots
library(loo) #for looic
library(patchwork) #to align multi-pane plots
library(grid) #to help patchwork with fonts
library(gsl) #for lambert_W0()
library(shinystan)
library(DT) #for making a nice html table 
source(here("R/functions.R"))
set.seed(123)

knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE,
                      echo = FALSE)
```
# Background  
This is the analysis supporting the paper see [Supplement 1](https://dylanmg.github.io/DL-CC-sockeye/output/html/S1_get_Smax_PR.html#Age_structure) for the calculation of $S_{MAX_{PR}}$ and [Supplement 2](https://dylanmg.github.io/DL-CC-sockeye/output/html/S2_diagnostics_figs.html#Posterior_distributions) for model diagnostics.  

### Load data  
Read in the PR data and the time series of stock recruit data then join into 1 big dataframe. $S_{MAX_{PR}}$ and error estimates are calculated in [Supplement 1](https://dylanmg.github.io/DL-CC-sockeye/output/html/S1_get_Smax_PR.html#Age_structure).   
```{r load-data}
SmaxPR <- read.table(here("output/data/SmaxPRs.txt"), 
                     header = TRUE) |>
  select(population, region, water_clarity, SmaxPR, SmaxPR_SD) |>
  dplyr::rename(pop = population) #to match below

SR_data <- read.table(here("data/FinalBroodSK.txt"), 
                      header = TRUE) |>
  select(-n.data, -n.stock) |> #dump useless vars
  mutate(watershed = gsub("_lake|_lakes", "", watershed)) |>
  #fix some names so it joins
  mutate(watershed = case_when(watershed == "owekino" ~ "owikeno",
                               watershed == "portjohn" ~ "port_john",
                               watershed == "freeda_brodie" ~ "freeda_lakes",
                               TRUE ~ watershed)) |> #leave the rest alone
  mutate(logRS = log(total_rec/exp_spawn)) |> #shorten names for plotting, etc.
  dplyr::rename(pop = watershed) |>
  relocate(pop, 1) |>
  filter(pop !="end_hill") |>
  arrange(pop, year) |>
  left_join(SmaxPR, by = "pop") |> #join it to have 1 object to reference
  mutate(pop = gsub("_", "-", pop)) |> #change delimiter for use later
  na.omit() |>
  group_by(pop) |>
  mutate(exp_spawn_scaled = scale(exp_spawn)) |>
  as.data.frame()

SR_pairs <- SR_data |>
  group_by(pop, region, water_clarity, SmaxPR, SmaxPR_SD) |>
  summarise(SR_pairs = n()) |>
  distinct() |>
  pull(SR_pairs)

pop_data <- read.csv(here("output/pub_tables/S1.csv")) |>
  dplyr::rename(pop = population) |>
  mutate(pop = gsub("_", "-", pop)) |> #change delimiter for use later
  mutate(pop_index = row_number()) |>
  cbind(SR_pairs)

colnames(pop_data) <- tolower(colnames(pop_data))

rm(SmaxPR, SR_pairs)
```
  
### Plot data  
Plot data that will be fit. Looking back at this may help get an idea why certain populations are difficult to fit.  

#### Raw data 
```{r SR-plots-raw}
#make helpers
max_facets <- 16
pages <- ceiling(length(unique(SR_data$pop))/max_facets)
j <- 1 #facet index helper

#make and print plots
for(i in 1:pages){
  sub_dat <- filter(SR_data, 
                    pop %in% unique(SR_data$pop)[j:(j+(max_facets-1))])
  
  p <- ggplot(sub_dat, 
              aes(exp_spawn/1000, total_rec)) +
    geom_point() +
    ggh4x::facet_wrap2(vars(pop), 
                       nrow=(max_facets/4), 
                       ncol=(max_facets/4), 
                       trim_blank=FALSE,
                       scales = "free") +
    expand_limits(x = 0, y = 0) +
    labs(title = paste0("Sockeye spawner-recruit relationship (", i, " of ", pages, ")"), 
         y = "Recruits", 
         x = "Spawners (thousands)") +
    theme_bw()
  
  print(p)
  
  j <- j + max_facets
}
```

#### Linear data 
```{r SR-plots-linear}
#make helpers
max_facets <- 16
pages <- ceiling(length(unique(SR_data$pop))/max_facets)
j <- 1 #facet index helper

#make and print plots
for(i in 1:pages){
  sub_dat <- filter(SR_data, 
                    pop %in% unique(SR_data$pop)[j:(j+(max_facets-1))])
  
  p <- ggplot(sub_dat, 
              aes(exp_spawn/1000, logRS)) +
    geom_point() +
    expand_limits(x = 0, y = 0) +
    ggh4x::facet_wrap2(vars(pop), 
                       nrow=(max_facets/4), 
                       ncol=(max_facets/4), 
                       trim_blank=FALSE,
                       scales = "free_x") +
    labs(title = paste0("Sockeye spawner-recruit relationship (", i, " of ", pages, ")"), 
         y = "ln(R/S)", 
         x = "Spawners (thousands)") +
    theme_bw()
  
  print(p)
  
  j <- j + max_facets
}
```

# Modelling  
We fit a variety of Ricker stock recruit relationships (i.e. variations of  $log(R/S) = log(\alpha) - \beta*S$) in a Bayesian framework with different priors and hierarchical structures ranging in complexity. In short, we compare different groupings of hierarchical $\alpha$ and uninformative/informative priors based on biogeoclimatic zones and photosynthetic rates. These will be fit in Stan via the `rstan` package.

Models took several forms:  
 
**1.** Independent models fit to each population with uninformative priors  
**2.** Uninformative $\beta$ priors fit to each population with uninformative hierarchical $\alpha$(s) where:  
\   \ **2a.** $\alpha$ is ungrouped and drawn from a distribution across all 69 populations OR  
\   \ **2b.** $\alpha$s categorically grouped by biogeoclimatic zone (n=3) and drawn from that group's distribution.  
**3.** Fits for each population with uninformative $\alpha$, and informative $\beta$ based on $S_{MAX_{PR}}$ estimates  
**4.** Informative, hierarchical $\alpha$(s) based on model performance in model 2 (i.e. grouped $\alpha$s or not), with informative priors on data based on $S_{MAX_{PR}}$.  

Models are checked in [Supplement 2](https://dylanmg.github.io/DL-CC-sockeye/output/html/S2_diagnostics_figs.html#Posterior_distributions) and Stan code is available in our [GitHub folder](https://github.com/DylanMG/DL-CC-sockeye/tree/main/Stan). 
```{r compile-stan-data}
logRS <- SR_data$logRS #response variable
pop <- as.numeric(as.factor(SR_data$pop)) #pop index vector
N <- as.numeric(length(logRS)) #n observations
J <- max(pop) #n populations
K <- max(as.numeric(as.factor(SR_data$region))) #n regions
region <- as.numeric(as.factor(SR_data$region)) #region index vector
reg_pop <- SR_data |> #region-pop key for indexing
  select(pop, region) |>
  distinct() |>
  pull(region) |>
  as.factor() |>
  as.numeric()

X <- matrix(rep(0, (N*J)), 
            nrow=N, 
            ncol=J) #empty predictor matrix to populate

for(i in 1:N){
  X[i, pop[i]] <- SR_data$exp_spawn[i] #populated!
}
int_X <- ifelse(X!=0, 1,0) #binary matrix to index when to "turn on" intercept 
SmaxPR <- as.numeric(unique(SR_data$SmaxPR)) #SmaxPR prior mus
SmaxPR_SD <- SR_data |> #SmaxPR SDs
  select(pop, SmaxPR_SD) |>
  distinct() |>
  pull(SmaxPR_SD) |> #got creative because a SD repeats
  as.numeric() 

stan_data <- list(logRS=logRS,
             N=N,
             pop=pop,
             J=J,
             K=K,
             region=region, 
             reg_pop=reg_pop,
             X=X,
             int_X=int_X,
             SmaxPR=SmaxPR,
             SmaxPR_SD=SmaxPR_SD)

rm(logRS, N, pop, J, K, region, reg_pop, X, int_X, SmaxPR, SmaxPR_SD)
```

```{r fits, results= "hide"}
refit <- FALSE #toggle if you want to refit models or used the stored model fits
iter <- 4000

if("m1" %in% list.files(here("output/model_fits")) & refit == FALSE){
  m1 <- readRDS(here("output/model_fits/m1"))
} else{
  m1 <- stan(here("Stan/model1.stan"), 
             data=stan_data, 
             model_name = "m1", 
             iter = iter)
  saveRDS(m1, file = here("output/model_fits/m1"))
}

if("m2a" %in% list.files(here("output/model_fits")) & refit == FALSE){
  m2a <- readRDS(here("output/model_fits/m2a"))
} else{
  m2a <- stan(here("Stan/model2a.stan"), 
              data=stan_data, 
              model_name = "m2a", 
              iter = iter) 
  saveRDS(m2a, file = here("output/model_fits/m2a"))
}

if("m2b" %in% list.files(here("output/model_fits")) & refit == FALSE){
  m2b <- readRDS(here("output/model_fits/m2b"))
} else{
  m2b <- stan(here("Stan/model2b.stan"), 
              data=stan_data, 
              model_name = "m2b", 
              iter = iter)
  saveRDS(m2b, file = here("output/model_fits/m2b"))
}

if("m3" %in% list.files(here("output/model_fits")) & refit == FALSE){
  m3 <- readRDS(here("output/model_fits/m3"))
} else{
  m3 <- stan(here("Stan/model3.stan"), 
             data=stan_data, 
             model_name = "m3", 
             iter = iter)
  saveRDS(m3, file = here("output/model_fits/m3"))
}

if("m4a" %in% list.files(here("output/model_fits")) & refit == FALSE){
  m4a <- readRDS(here("output/model_fits/m4a"))
} else{
  m4a <- stan(here("Stan/model4a.stan"), 
             data=stan_data, 
             model_name = "m4a",  
             iter = iter)
  saveRDS(m4a, file = here("output/model_fits/m4a"))
}

if("m4b" %in% list.files(here("output/model_fits")) & refit == FALSE){
  m4b <- readRDS(here("output/model_fits/m4b"))
} else{
  m4b <- stan(here("Stan/model4b.stan"), 
             data=stan_data, 
             model_name = "m4b",  
             iter = iter)
  saveRDS(m4b, file = here("output/model_fits/m4b"))
}

models <- list(m1=m1, 
               m2a=m2a, 
               m2b=m2b, 
               m3=m3, 
               m4a=m4a, 
               m4b=m4b)
#get some data for figs ready
all_draws <- get_draws(models, pop_data)
```

# Tables  

Table 1. looic among models  
```{r table1}
mod_notes <- data.frame(Model = c("m1", "m2a", "m2b", "m3", "m4a", "m4b"), 
                        Description = c("status quo", "hierarchical prod.", "regional-hierarchical prod.", 
                                        "informative capacity only", "hierarcical prod. & informative capacity", 
                                        "regional-hierarcical prod. & informative capacity"),
                        Note = c("max. R-hat = 1.09, low min. ESS", 
                                 "low min. ESS", 
                                 "4 div. transitions, max. R-hat = 1.14, low min. ESS",
                                 "low min. ESS",
                                 "",
                                 "35 div. transitions"))

OOS_loo_table <- read.csv(here("output/CV/loos_table.csv")) |>
  filter(Metric =="looic") |>
  rename(LOOIC = Estimate) |>
  mutate(LOOIC = round(LOOIC, 1), 
         SE = round(SE, 1)) |> 
  left_join(mod_notes, by = "Model") |>
    select(Model, Description, Weight, LOOIC, SE, Note) |>
    arrange(desc(Weight))

datatable(OOS_loo_table,
          rownames = FALSE,
          options = list(dom = 't'))
write.csv(OOS_loo_table, here("output/pub_tables/table1.csv"), row.names = FALSE)
```

# Figures  

**add S/Smsy fig**

```{r SSmsy}
#get median Smsy from model 4 for Will 
m2b_Smsy <- all_draws |>
  filter(model == "m2b") |>
  group_by(pop) |>
  summarise(Smsy_med = median(Smsy))
#write.csv(m2b_Smsy, "C:/Users/GLASERD/Desktop/m2b_Smsy_med.csv", row.names = FALSE)

S_Smsy <- read.csv(here("data/M4_Smsy_med+recent escapements_DG.csv")) |>
  mutate(s_smsy = as.numeric(s_smsy))

length(which(is.na(S_Smsy$s_smsy)))/nrow(S_Smsy) #%NAs
length(which(S_Smsy$s_smsy>1))/length(which(!is.na(S_Smsy$s_smsy))) #% above Smsy
length(which(S_Smsy$s_smsy<1))/length(which(!is.na(S_Smsy$s_smsy))) #% below Smsy
```
### Figure 1  

```{r fig1, fig.width=5.62, fig.height=4.215, dpi= 300}
fig_1_data <- filter(all_draws, pop %in% c("Johnston", "Kainet", "Slamgeesh", "Skidegate", 
                                            "Mathers", "Morice", "Atnarko", "Canoona", "Owikeno")) |>
  left_join(pop_data, by = "pop") |>
  mutate(data_quality = case_when(pop %in% c("Mathers", "Johnston", "Slamgeesh") ~ "low", #<15,  
                                  pop %in% c("Atnarko", "Canoona", "Kainet") ~ "medium", #20-40
         pop %in% c("Morice", "Owikeno", "Skidegate") ~ "high"),#40+
         data_quality = factor(data_quality, levels = c("low", "medium", "high"))) 
  
fig_1_data |>
  select(pop, region.x, data_quality, sr_pairs) |>
  distinct() |>
  arrange(region.x, data_quality)

p <- ggplot(fig_1_data, aes(model, log_a)) +
  geom_violin(fill = "grey",
              draw_quantiles = 0.5) +
  facet_grid(region.x~data_quality) +
  geom_label(x=2.4, y = 3.7, aes(label = pop), stat = "unique", family = "serif")+ #writes it a ton
  labs(x = "Model", y = expression("log("*alpha*")")) +
  theme_bw() +
  theme(text = element_text(family="serif"))

p

ggsave(here("output/pub_figs/fig1.png"), 
       plot = p,
       width = 5.61, 
       height = 4.22, 
       dpi = 300)
```
### Figure 2   

```{r fig2 wrangle}
index_df <- cbind(pop_data$pop, as.numeric(as.factor(pop_data$pop))) #helper df for the loop 

lakes <- c("Johnston", "Kainet", "Owikeno", "Slamgeesh", "Atnarko", "Morice", "Mathers", 
           "Canoona", "Skidegate")

fig2_preds <- NULL
fig2_data <- NULL

for(i in unique(lakes)){
  for(j in c("m1", "m2b")){
    sub_draw <- filter(all_draws, pop==i, model==j)
    sub_dat <- filter(SR_data, pop==tolower(i))
    
    max_samples <- nrow(sub_draw)
    spw <- seq(0, max(sub_dat$exp_spawn), length.out = 100)
    SR_pred <- matrix(NA, 100, max_samples)
    
    for(k in 1:max_samples){
      r <- sample(seq(1,max_samples), 1, replace=TRUE)
      log_a <- sub_draw$log_a[r]
      b <- sub_draw$b[r]
      SR_pred[,k] <- (exp(log_a)*spw*exp(-b*spw))
    }
    SR_pred <- cbind(spw,t(apply(SR_pred,c(1),quantile,probs=c(0.1,0.5,0.9),na.rm=T))) |>
      as.data.frame() |>
      round(2) |>
      mutate(pop = i, 
             model = j)
    
    fig2_preds <- bind_rows(SR_pred, fig2_preds)
    
    fig2_data <- bind_rows(sub_dat, fig2_data)
  }
}

colnames(fig2_preds) <- c("Spawn", "Rec_lwr", "Rec_med", "Rec_upr", "pop", "model")

fig2_preds <- fig2_preds |>
  left_join(pop_data, by = "pop") |>
  mutate(pop = factor(str_to_title(pop), levels = lakes)) |>
    mutate(data_quality = case_when(pop %in% c("Mathers", "Johnston", "Slamgeesh") ~ "low", #<15,  
                                  pop %in% c("Atnarko", "Canoona", "Kainet") ~ "medium", #20-40
         pop %in% c("Morice", "Owikeno", "Skidegate") ~ "high"),#40+
         data_quality = factor(data_quality, levels = c("low", "medium", "high"))) 

fig2_data <- fig2_data |>
  left_join(pop_data, by = "pop") |>
    mutate(pop = factor(str_to_title(pop), levels = lakes)) |>

  mutate(data_quality = case_when(pop %in% c("Mathers", "Johnston", "Slamgeesh") ~ "low", #<15,  
                                  pop %in% c("Atnarko", "Canoona", "Kainet") ~ "medium", #20-40
         pop %in% c("Morice", "Owikeno", "Skidegate") ~ "high"),#40+
         data_quality = factor(data_quality, levels = c("low", "medium", "high"))) 

p <- ggplot() +
  geom_ribbon(data = fig2_preds, aes(x = Spawn/1000, ymin = Rec_lwr/1000, ymax = Rec_upr/1000, 
                                      fill = model, color = model), alpha=0.5) +
  geom_line(data = fig2_preds, aes(x = Spawn/1000, y = Rec_med/1000, color=model), linewidth = 1) +
  geom_point(data = fig2_data, aes(x=exp_spawn/1000, y=total_rec/1000)) +
  #ggh4x::facet_grid2(region.x~data_quality, scales = "free", independent=c("x", "y")) +
  facet_wrap(~pop, scales = "free") +
  theme_classic() +
  scale_fill_viridis_d(name = "Model", 
                       labels = c("1", "2b")) +
  scale_color_viridis_d() +
  theme(text = element_text(family="serif"),
        legend.position= "bottom", 
        legend.title = element_text(size=9),
        legend.text = element_text(size=7),
        legend.key.size = unit(.5, "cm")) +
        labs(y = "Recruits (1000s)", x = "Spawners (1000s)") +
  scale_x_continuous(breaks = pretty_breaks()) +
  guides(color="none")

p

ggsave(here("output/pub_figs/fig2.png"), 
       plot = p,
       width = 5.61, 
       height = 4.22, 
       dpi = 300)
```

### Figure 3  

```{r fig3, fig.width=5.62, fig.height=4.215, dpi= 300}
p <- ggplot(filter(all_draws, model == "m2b"), aes(region, log_a)) +
  geom_violin(fill = "grey",
              draw_quantiles = 0.5) +
  labs(title = "Regional Productuvity in Model 2b", 
        y = expression("log("*alpha*")"), 
       x = "Region") +
  theme_bw() +
  scale_x_discrete(labels = c("coastal fjord", "interior", "low coastal")) +
  theme(legend.position = "bottom", 
        text = element_text(family="serif")) 
p
ggsave(here("output/pub_figs/fig3.png"), 
       plot = p,
       width = 5.61, 
       height = 4.22, 
       dpi = 300)

print(p)
```

### Figure 4  
Difference in LOOIC weight

```{r fig4, fig.width=5.62, fig.height=4.215, dpi= 300}
pop_weights_quo <- read.csv(here("output/CV/pop_weights_quo.csv"))

fig4_data <- pop_weights_quo |>
  pivot_wider(names_from = Model, values_from = Weight) |>
  mutate(pop = str_to_title(gsub("-", " ", Pop))) |>
  left_join(pop_data, by = "pop") |>
  select(-Pop) |>
  relocate(pop, 1)

p1 <- ggplot(fig4_data,aes(sr_pairs, m2b)) +
  theme_classic() +
  geom_point(aes(color = region)) +
  geom_hline(yintercept = 0.5, lty = 2) +
  annotate("text", x = 5, y = 1.2, label = "(A)") +
  theme(legend.position = "bottom", 
        legend.title = element_text(size = 9),
        legend.text = element_text(size = 7),
        legend.key.size = unit(.1, "cm"),
        text = element_text(family = "serif")) + 
  scale_color_viridis_d(name = "Region", 
                       labels = c("coastal fjord", "interior", "low coastal")) +
  stat_smooth(method = "loess", 
              color = "black",
              se = FALSE, 
              alpha = 0.5) +
  labs(x = "Years of SR pairs", 
       y = NULL)

p2 <- ggplot(filter(fig4_data, est_origin != "both"), aes(est_origin, m2b)) +
  geom_violin(draw_quantiles = 0.5) +
  geom_jitter(width = 0.25) +
  geom_hline(yintercept = 0.5, lty = 2) +
  annotate("text", x= .52, y = 1.2, label = "(B)") +
  theme_classic() +
  theme(text = element_text(family="serif"),
        legend.position= "bottom", 
        legend.title = element_text(size=9),
        legend.text = element_text(size=7),
        legend.key.size = unit(.5, "cm"),) + 
  scale_fill_viridis_d(name = "Estimate method", 
                       labels = c("empirical", "modeled")) +
  scale_x_discrete(labels = c("empirical", "modeled")) +
  labs(x= "PR estimate method", y=NULL) 

p3 <- ggplot(fig4_data, aes(region, m2b)) +
      geom_violin(#aes(fill=est_origin), 
                  alpha = 0.5, 
                  draw_quantiles = 0.5) +
  geom_jitter(width = 0.25) +
  geom_hline(yintercept = 0.5, lty = 2) +
  annotate("text", x= .6, y = 1.2, label = "(C)") +
  theme_classic() +
  theme(text = element_text(family="serif")) +
  scale_fill_viridis_d() +
  scale_x_discrete(labels = c("coastal fjord", "interior", "low coastal")) +
  labs(x= "Region", y=NULL) +
  guides(fill="none")

p4 <- ggplot(fig4_data, aes(water_clarity, m2b)) +
  geom_violin(#aes(fill=est_origin), 
              alpha = 0.5, 
              draw_quantiles = 0.5) +
  geom_jitter(width = 0.25) +
  geom_hline(yintercept = 0.5, lty = 2) +
  annotate("text", x= .6, y = 1.2, label = "(D)") +
  theme_classic() +
  theme(text = element_text(family="serif")) + 
  scale_fill_viridis_d() +
  labs(x= "Water clarity", y=NULL) +
  guides(fill="none")

pw <- ((p1/p3)|(p2/p4))

p <- gridExtra::grid.arrange(patchworkGrob(pw), 
                        left = textGrob(expression("M2b looic stacked weight (m2b vs m1)"), 
                                        gp = gpar(fontfamily = "serif"), rot=90))
#^ from https://statisticsglobe.com/add-subscript-and-superscript-to-plot-in-r example 3

print(p)
ggsave(here("output/pub_figs/fig4.png"), 
       plot = p,
       width = 5.61, 
       height = 4.22, 
       dpi = 300)
```



