---
title: "Combining hierarchical models and habitat data improves assessment of data limited salmon stocks "
subtitle:  |
    | Marine and Coastal Fisheries
    | Supplement 2: Model diagnostics and reporting full posteriors 
author: "W.I. Atlas, D.M. Glaser, B.M. Connors, C.A. Holt, D.T. Selbie, S. Cox-Rogers, C. Carr-Harris, E. Hertz, and J.W. Moore."
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output/html") })
output: 
  bookdown::html_document2:
    theme: cerulean
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: false
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: sentence
#bibliography: references.bib
#csl: "https://raw.githubusercontent.com/citation-style-language/styles/master/canadian-journal-of-fisheries-and-aquatic-sciences.csl"
link-citations: true
always_allow_html: yes
---

```{r pkg-setup, echo = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
library(here)
library(ggh4x) #to control facetting
library(rstan) #to interface R with Stan
library(bayesplot) #for mcmc_combo plots
library(loo) #for looic
library(patchwork) #to align multi-pane plots
library(grid) #to help patchwork with fonts
library(gsl) #for lambert_W0()
library(DT) #for making a nice html table 
library(shinystan)
source(here("R/functions.R"))

knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE,
                      echo = FALSE,
                      fig.path = paste0(here("output/plots"), "/"))
```
# Background  
These are some model diagnostics and posteriors from the spawner-recruit analyses in the article mentioned above.
```{r load fits and data}
models <- c("m1", "m2a", "m2b", "m3", "m4")

#use map to read in all model fits as list
fits <- map(models, function(x){
  readRDS(file = here(paste0("output/model_fits/", x)))
})

SmaxPR <- read.table(here("output/data/SmaxPRs.txt"), 
                     header = TRUE) |>
  select(population, region, water_clarity, SmaxPR, SmaxPR_SD) |>
  dplyr::rename(pop = population) #to match below

SR_data <- read.table(here("data/FinalBroodSK.txt"), 
                      header = TRUE) |>
  select(-n.data, -n.stock) |> #dump useless vars
  mutate(watershed = gsub("_lake|_lakes", "", watershed)) |>
  #fix some names so it joins
  mutate(watershed = case_when(watershed == "owekino" ~ "owikeno",
                               watershed == "portjohn" ~ "port_john",
                               watershed == "freeda_brodie" ~ "freeda_lakes",
                               TRUE ~ watershed)) |> #leave the rest alone
  mutate(logRS = log(total_rec/exp_spawn)) |> #shorten names for plotting, etc.
  dplyr::rename(pop = watershed) |>
  relocate(pop, 1) |>
  filter(pop !="end_hill") |>
  arrange(pop, year) |>
  left_join(SmaxPR, by = "pop") |> #join it to have 1 object to reference
  mutate(pop = gsub("_", "-", pop)) |> #change delimiter for use later
  na.omit() #drops years missing logRS

SR_pairs <- SR_data |>
  group_by(pop, region, water_clarity, SmaxPR, SmaxPR_SD) |>
  summarise(SR_pairs = n()) |>
  distinct() |>
  pull(SR_pairs)

pop_data <- read.csv(here("output/pub_tables/S1.csv")) |>
  dplyr::rename(pop = population) |>
  mutate(pop = gsub("_", "-", pop)) |> #change delimiter for use later
  mutate(pop_index = row_number()) |>
  cbind(SR_pairs)

colnames(pop_data) <- tolower(colnames(pop_data))
```

## Diagnostics  
Here we check if chains have mixed, the maximum $\hat{R}$ is less than 1.05, and that the effective sample size is at least 10% of the iterations, or in our case 400. 

### Model 1  
Fits for each population with uninformative priors. We won't plot any `mcmc_combo()` for this because there's too many parameters (3 per population, since all are fit independently) to nicely visualize in a summary doc. You can fork the [code](https://github.com/DylanMG/sockeye-hab-SRR-Stan/blob/main/R/model_fitting_comparison.Rmd) and run `rstanarm::launch_shinystan()` to check  models if you want. Just look at the sigma as a check, then report max Rhats later.  
```{r m1-diag}
mcmc_combo(fits[[1]], combo = c("dens_overlay", "trace"), 
           pars = c("sigma"), 
           gg_theme = legend_none())
#plot(fits[[1]])
fit1.summary  <- as.data.frame(rstan::summary(fits[[1]])$summary)
paste("max R-hat =", max(fit1.summary$Rhat, na.rm = TRUE))
paste("min ESS =", min(fit1.summary$n_eff, na.rm = TRUE)) #WHAT IS GOING ON WITH NEFF HERE!! max is more than iter

ggplot(fit1.summary, aes(n_eff)) +
  geom_histogram() +
  geom_vline(xintercept = 400, lty=2) + #could annotate a "10%" here too 
  theme_classic() +
  labs(x = "effective sample size")
```

### Model 2a  
Uninformative hierarchical $\alpha$  

```{r m2a-diag}
mcmc_combo(fits[[2]], combo = c("dens_overlay", "trace"), 
           pars = c("log_a", "sd_a_pop", "sigma"), 
           gg_theme = legend_none())
fit2a.summary  <- as.data.frame(rstan::summary(fits[[2]])$summary)
paste("max R-hat =", max(fit2a.summary$Rhat, na.rm = TRUE))
paste("min ESS =", min(fit2a.summary$n_eff, na.rm = TRUE))
#loo(fits[[2]])
#plot(loo(fits[[2]]))
ggplot(fit2a.summary, aes(n_eff)) +
  geom_histogram() +
  geom_vline(xintercept = 400, lty=2) +
  theme_classic() +
  labs(x = "effective sample size")

#plot PPC
logRS <- SR_data$logRS
model.pars2a <- rstan::extract(fits[[2]])
logRS_rep <- model.pars2a$logRS_rep

#ppc_dens_overlay(logRS, logRS_rep) +
#  theme(legend.position = "none") +
#  labs(y = "density", x = "y_est")
```

```{r m2a-variants}
# further diagnosing divergent transitions
if(FALSE){
posterior_2a <- as.array(fits[[2]])
np_2a <- nuts_params(fits[[2]])
#mcmc_parcoord(posterior_2a, np = np_2a)

#launch_shinystan(fits[[2]])
fit2a.HN <- readRDS(here("output/model_fits/alternates/m2a_HN"))
fit2a.UC <- readRDS(here("output/model_fits/alternates/m2a_uncentered"))
fit2a.UC.HN <- readRDS(here("output/model_fits/alternates/m2a_uncentered_HN"))
fit4.HN <- readRDS(here("output/model_fits/alternates/m4_HN"))

fit2a.HN.summary  <- as.data.frame(rstan::summary(fit2a.HN)$summary)
fit2a.UC.summary  <- as.data.frame(rstan::summary(fit2a.UC)$summary)
fit2a.UC.HN.summary  <- as.data.frame(rstan::summary(fit2a.UC.HN)$summary)
fit4.HN.summary  <- as.data.frame(rstan::summary(fit4.HN)$summary)

#launch_shinystan(fit2a.HN)
#launch_shinystan(fit2a.UC)
#launch_shinystan(fit2a.UC.HN)
#launch_shinystan(fit4.HN)
}

```

### Model 2b  
Now we make $\alpha$ hierarchical to each biogeoclimatic zone; each of the 3 biogeoclimatic regions gets it *own* distribution that population level random effects are drawn from.  
```{r m2b-diag}
mcmc_combo(fits[[3]], combo = c("dens_overlay", "trace"), 
           pars = c("log_a", "sd_a_pop", "sd_a_region", "sigma"), 
           gg_theme = legend_none())
mcmc_combo(fits[[3]], combo = c("dens_overlay", "trace"), 
           pars = c("log_a_region[1]", "log_a_region[2]", "log_a_region[3]"), 
           gg_theme = legend_none()) 
fit2b.summary  <- as.data.frame(rstan::summary(fits[[3]])$summary)
paste("max R-hat =", max(fit2b.summary$Rhat, na.rm = TRUE))
paste("min ESS =", min(fit2b.summary$n_eff, na.rm = TRUE))
#loo(fits[[3]])
#plot(loo(fits[[3]]))
ggplot(fit2b.summary, aes(n_eff)) +
  geom_histogram() +
  geom_vline(xintercept = 400, lty=2) +
  theme_classic() +
  labs(x = "effective sample size")
```

### Model 3  
Models fit to each lake with an informative $\beta$ prior based on $S_{MAX_{PR}}$, and an uninformative prior for $\alpha$. We'll just plot the sigma here since everything else is independent. 
```{r m3-diag}
#need to find a better way to check the chains?
mcmc_combo(fits[[4]], combo = c("dens_overlay", "trace"), 
           pars = c("sigma"), 
           gg_theme = legend_none())
fit3.summary  <- as.data.frame(rstan::summary(fits[[4]])$summary)
paste("max R-hat =", max(fit3.summary$Rhat, na.rm = TRUE))
paste("min ESS =", min(fit3.summary$n_eff, na.rm = TRUE))
#loo(fits[[4]])
#plot(loo(fits[[4]]))
ggplot(fit3.summary, aes(n_eff)) +
  geom_histogram() +
  geom_vline(xintercept = 400, lty=2) +
  theme_classic() +
  labs(x = "effective sample size")
```

### Model 4
Informative, hierarchical $\alpha$ based on model performance in model 2 (i.e. grouped $\alpha$ or not), with informative priors on data based on Smax_PR.   
```{r m4-diag}
mcmc_combo(fits[[5]], combo = c("dens_overlay", "trace"), 
           pars = c("log_a", "sd_a_pop", "sigma"), 
           gg_theme = legend_none())
mcmc_combo(fits[[5]], combo = c("dens_overlay", "trace"), 
           pars = c("log_a_region[1]", "log_a_region[2]", "log_a_region[3]"), 
           gg_theme = legend_none()) 
fit4.summary  <- as.data.frame(rstan::summary(fits[[5]])$summary)
paste("max R-hat =", max(fit4.summary$Rhat, na.rm = TRUE))
paste("min ESS =", min(fit4.summary$n_eff, na.rm = TRUE))
#loo(fits[[5]])
#plot(loo(fits[[5]]))
ggplot(fit4.summary, aes(n_eff)) +
  geom_histogram() +
  geom_vline(xintercept = 400, lty=2) +
  theme_classic() +
  labs(x = "effective sample size")
```

```{r get draws}
#OPTIONAL TO SHOW
#After extracting draws we can look at regional productivity by model.
#get some data for figs ready
all_draws <- get_draws(fits, pop_data)

#then get regional productivity
reg_prod <- all_draws |>
  group_by(model, region) |>
  summarise(log_a_mu = mean(log_a),
            log_a_sd = sd(log_a), 
            a_mu = round(exp(log_a_mu), 2), 
            a_sd = round(exp(log_a_sd), 2)) |>
  mutate(log_a_mu = round(log_a_mu, 2), 
         log_a_sd = round(log_a_sd, 2))

#datatable(reg_prod, rownames = FALSE, 
#          caption = NA,
#          filter = 'top', 
#          options = list(pageLength = 15, autoWidth = TRUE))

#get median Smsy from model 4 for Will 
M4_Smsy <- all_draws |>
  filter(model == "m4") |>
  group_by(pop) |>
  summarise(Smsy_med = median(Smsy))

write.csv(M4_Smsy, here("output/data/M4_Smsy_med.csv"), row.names = FALSE)
```

## Posterior distributions  

### $S_{MSY}$ posteriors  
```{r Smsy plot}
max_facets <- 16
pages <- ceiling(length(unique(SR_data$pop))/max_facets)

j <- 1 #facet index helper

#make and print plots
for(i in 1:pages){
  sub_dat <- filter(all_draws, 
                    pop %in% gsub("-", " ", 
                                  str_to_title(unique(SR_data$pop)))[j:(j+(max_facets-1))])
  
  p <- ggplot(sub_dat, 
              aes(model, log(Smsy))) +
    geom_violin(fill = "grey", 
                draw_quantiles = 0.5) +
    ggh4x::facet_wrap2(vars(pop), 
                       nrow=(max_facets/4), 
                       ncol=(max_facets/4), 
                       trim_blank=FALSE,
                       scales = "free_y") +
    labs(title = paste0("Smsy estimate (", i, " of ", pages, ")"), 
         y = "log(Spawners)", 
         x = "model") +
    theme_bw() +
  guides(fill="none")
  
  print(p)
  
  j <- j + max_facets
}
```

### log($\alpha$) posteriors
```{r a plots}
j <- 1 #facet index helper

#make and print plots
for(i in 1:pages){
  sub_dat <- filter(all_draws, 
                    pop %in% gsub("-", " ", 
                                  str_to_title(unique(SR_data$pop)))[j:(j+(max_facets-1))])
  
  p <- ggplot(sub_dat, 
              aes(model, log_a)) +
    geom_violin(fill = "grey",
                draw_quantiles = 0.5) +
    ggh4x::facet_wrap2(vars(pop), 
                       nrow=(max_facets/4), 
                       ncol=(max_facets/4), 
                       trim_blank=FALSE,
                       scales = "free_y") +
    labs(title = paste0("log(a) estimate (", i, " of ", pages, ")"), 
         y = "log_alpha", 
         x = "model") +
    theme_bw()+
  guides(fill="none")
  
  print(p)
  
  j <- j + max_facets
}
```

### $\beta$ posteriors  
```{r b plots}
j <- 1 #facet index helper

#make and print plots
for(i in 1:pages){
  sub_dat <- filter(all_draws, 
                    pop %in% gsub("-", " ", 
                                  str_to_title(unique(SR_data$pop)))[j:(j+(max_facets-1))])
  
  p <- ggplot(sub_dat, 
              aes(model, b)) +
    geom_violin(fill = "grey",
              draw_quantiles = 0.5) +
    ggh4x::facet_wrap2(vars(pop), 
                       nrow=(max_facets/4), 
                       ncol=(max_facets/4), 
                       trim_blank=FALSE,
                       scales = "free_y") +
    scale_fill_viridis_d() +
    labs(title = paste0("beta estimate (", i, " of ", pages, ")"), 
         y = "beta", 
         x = "model") +
    theme_bw()+
  guides(fill="none")
  
  print(p)
  
  j <- j + max_facets
}
```

### Models 1 and 4 fit to data

```{r m1 & m4 to data}
fig_preds <- NULL
fig_data <- NULL

for(i in unique(all_draws$pop)){
  for(j in c("m1", "m4")){ #add or adjust the models as seen fit
    sub_draw <- filter(all_draws, pop==i, model==j)
    sub_dat <- filter(SR_data, pop==gsub(" ", "-", tolower(i)))
    
    max_samples <- nrow(sub_draw)
    spw <- seq(0, max(sub_dat$exp_spawn), length.out = 100)
    SR_pred <- matrix(NA, 100, max_samples)
    
    for(k in 1:max_samples){
      r <- sample(seq(1,max_samples), 1, replace=TRUE)
      log_a <- sub_draw$log_a[r]
      b <- sub_draw$b[r]
      SR_pred[,k] <- (exp(log_a)*spw*exp(-b*spw))
    }
    SR_pred <- cbind(spw,t(apply(SR_pred,c(1),quantile,probs=c(0.05,0.5,0.95),na.rm=T))) |>
      as.data.frame() |>
      round(2) |>
      mutate(pop = i, 
             model = j)
    
    fig_preds <- bind_rows(fig_preds, SR_pred)
    
    fig_data <- bind_rows(fig_data, sub_dat)
  }
}

colnames(fig_preds) <- c("Spawn", "Rec_lwr", "Rec_med", "Rec_upr", "pop", "model")

fig_data <- mutate(fig_data, pop = gsub("-", " ", str_to_title(pop))) 

j <- 1
max_facets <- 8

pages <- ceiling(length(unique(SR_data$pop))/max_facets)
for(i in 1:pages){
  sub_data <- filter(fig_data, 
                    pop %in% unique(fig_preds$pop)[j:(j+(max_facets-1))])

  sub_preds <- filter(fig_preds, 
                    pop %in% unique(fig_preds$pop)[j:(j+(max_facets-1))])
  
  p <- ggplot() +
    geom_ribbon(data = sub_preds, aes(x = Spawn/1000, ymin = Rec_lwr/1000, ymax = Rec_upr/1000, 
                                        fill = model, color = model), alpha=0.5) +
    geom_line(data = sub_preds, aes(x = Spawn/1000, y = Rec_med/1000, color=model), size = 1) +
    geom_point(data = sub_data, aes(x=exp_spawn/1000, y=total_rec/1000)) +
    ggh4x::facet_wrap2(vars(pop), 
                       nrow=2, 
                       ncol=4, 
                       trim_blank=FALSE, 
                       scales = "free") +
    theme_classic() +
    scale_fill_viridis_d() +
    scale_color_viridis_d() +
    theme(text = element_text(family="serif"),
          legend.position= "bottom", 
          legend.title = element_text(size=9),
          legend.text = element_text(size=7),
          legend.key.size = unit(.5, "cm")) +
        labs(title = paste0("S-R models m1 & m4 (", i, " of ", pages, ")"), 
         y = "Recruits (thousands)", 
         x = "Spawners (thousands)")

  print(p)
  j <- j + max_facets
}
```